---
title: "Camtrap Meta"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Camtrap synthesis 
1. Does sampling effort predict animal abundances or diversity for a system?  
2. Is there a sweet spot or threshold in sampling?


### Data 
[Measures for Event Counts](https://www.rdocumentation.org/packages/metafor/versions/2.4-0/topics/escalc)  
Various measures can be used to characterize individual groups when the dependent variable assessed is an event count. Here, one needs to specify xi and ti, denoting the number of events that occurred and the total person-times at risk, respectively. The options for the measure argument are then:

"IR" for the raw incidence rate.

"IRLN" for the log transformed incidence rate.

"IRS" for the square root transformed incidence rate.

"IRFT" for the Freeman-Tukey transformed incidence rate (Freeman & Tukey, 1950).

Studies with zero events can be problematic, especially for the log transformed incidence rate. Adding a small constant to the number of events is a common solution to this problem. When to="only0" (the default), the value of add (the default is 1/2; but see ‘Note’) is added to xi only in the studies that have zero events. When to="all", the value of add is added to xi in all studies. When to="if0all", the value of add is added to xi in all studies, but only when there is at least one study with zero events. Setting to="none" or add=0 has the same effect: No adjustment to the observed number of events is made. Depending on the outcome measure and the data, this may lead to division by zero inside of the function (when this occurs, the resulting value is recoded to NA).  

Also see Cochrane handbook [ch 6.7](https://training.cochrane.org/handbook/current/chapter-06#_Ref523230081)


```{r}
library(tidyverse)
data <- read_csv("C:/Users/Nargol Ghazian/Desktop/PhD-Shelter-Shrub-Climate/Camtrap Review/review.csv") %>% 
  filter(days < 50000 & captures < 80000) %>% 
  rename(n_cams = 'number cams') %>% 
  mutate(n_cams = as.numeric(n_cams)) %>% 
  filter(bait == "no" & ecosystem != "aquatic")#clean up and filter data
data

library(metafor)
#abundance of animals
data_e1 <- escalc(measure = "IR", xi = captures, ti = days, ni = n_cams,  data = data) %>% 
  filter(yi != "NA")

#xi=variable of interest that's count data
#ni=vector to specify the sample/group sizes
#ti= vector to specify the total person-times
#n.animals/n.cams/n.days


#animal richness
data_e2 <- escalc(measure = "IR", xi = richness, ti = days, ni= n_cams, data = data) %>% 
  filter(yi != "NA")
#number of species/n.cams/n.days


#animal captures without days in the effect size for regression
data_e3 <- escalc(measure = "IR", xi = captures, ti = n_cams,  data = data) %>% 
  filter(yi != "NA")

#richness wihtout days in the effect size for regression
data_e4 <- escalc (measure = "IR", xi = richness, ti = n_cams, data = data)

#join all the data that has days in the effect size index to be used for the forest plots as sampling effort
data_1 <- data_e1 %>% 
  rename(capture_rate = yi, var_caprate = vi)

data_2 <- data_e2 %>% 
  rename(richness_rate = yi, var_richnessrate = vi) %>% 
  select(ID, richness_rate, var_richnessrate)

data_all <- left_join(data_1, data_2, by = "ID")
write.csv(data_all, "C:/Users/Nargol Ghazian/Desktop/PhD-Shelter-Shrub-Climate/Camtrap Review/Final Data with Effect Sizes.csv", row.names = FALSE)
#write.csv save complete dataset 


summary_data <- data_all %>% 
  group_by(ecosystem) %>% 
  summarise(n_cams = mean(n_cams), capture_rate = mean(capture_rate), richness_rate = mean(richness_rate, na.rm = TRUE), var_caprate= sd(var_caprate, na.rm=TRUE), var_richnessrate = sd(var_richnessrate, na.rm = TRUE))#filtered out mixed forest

```

### Publication Viz

```{r}

#FIGURE 1 A AND B

ggplot(data_e3, aes(days, yi)) +
  geom_point(aes(color = ecosystem)) + 
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 2, color = "black", se = FALSE) +
  labs(x = "Number of Days", y = "Number of Animals/Number of Cameras")+ theme_classic()+ labs(colour = "Ecosystem")

ggplot(data_e4, aes(days, yi, weight = vi)) +
  geom_point(aes(color = ecosystem)) + 
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 2, color = "black", size = 2, se = FALSE) +
  labs(x = "Number of Days", y = "Number of Species/Number of Cameras")+ theme_classic()+ labs(colour = "Ecosystem")


#FIGURE 2 A AND B

ggplot(summary_data, aes(ecosystem, richness_rate, color = factor(ecosystem))) +
  geom_point(size =2) +
  geom_errorbar(aes(ymin = richness_rate - var_richnessrate, ymax = richness_rate + var_richnessrate), size= 0.7, width=0.1, position = position_dodge(width = 0.5)) +
  coord_flip() +
  geom_hline(yintercept = 1, colour="grey", linetype = "longdash", size = 2) + 
  labs(x = "Ecosystem", y = "Number of Species/Number of Cameras/Day") + theme_classic()+ theme(legend.position = "none")

ggplot(summary_data, aes(ecosystem, capture_rate, color = factor(ecosystem))) +
  geom_point(size =2) +
  geom_errorbar(aes(ymin = capture_rate - var_caprate, ymax = capture_rate + var_caprate), size= 0.5, width=0.1, position = position_dodge(width = 0.5)) +
  coord_flip() +
  geom_hline(yintercept = 1, colour="grey", linetype = "longdash", size = 2) + 
  labs(x = "Ecosystem", y = "Number of Animals/Number of Cameras/Day") + theme_classic()+ theme(legend.position = "none")#variance is too small to see

```








### Meta models
Do a meta weighting captures by number of cameras over days. And richness by number of cams.

```{r}

library(broom)
library(tidymodels)

#models for the forest plots

#x = captures
#y = number of cams
#y = days

mod1.1 <- rma(yi = yi, sei = vi, method = "ML", mods = ~ecosystem -1, test = "knha", control=list(stepadj=0.5), data = data_e1)
summary(mod1.1)
m1.1 <- tidy(mod1.1)
m1.1
plot(mod1.1)


#x = richness
#y = number of cams
#y = days
mod1.2 <- rma(yi = yi, sei = vi, method = "ML", mods = ~ecosystem -1, test = "knha", control=list(stepadj=0.5), data = data_e2)#using ecosystem as a moderator
summary(mod1.2)
m1.2 <- tidy(mod1.2)
m1.2
plot(mod1.2)


#regression models with days as a moderator

#abundance
mod2.1 <- rma(yi = yi, sei = vi, method = "ML", mods = ~days, test = "knha", control=list(stepadj=0.5), data = data_e3)
summary(mod2.1)
m2.1 <- tidy(mod2.1)
m2.1
plot(mod2.1)

#richness
mod2.2 <- rma(yi = yi, sei = vi, method = "ML", mods = ~days, test = "knha", control=list(stepadj=0.5), data = data_e4)
summary(mod2.2)
m2.2 <- tidy(mod2.2)
m2.2
plot (mod2.2)


```


## Viz  
Need to weight points but sampling effort - n_cams/number of days for capture rate and richness. Then you can have an index that only has richness/n_cams or abundance/n_camps and regress that against days. 

```{r}
#forest plots for indeces with days as a measure of sampling effort

ggplot(summary_data, aes(ecosystem, richness_rate, color = factor(ecosystem))) +
  geom_point(size =2) +
  geom_errorbar(aes(ymin = richness_rate - var_richnessrate, ymax = richness_rate + var_richnessrate), size= 0.7, width=0.1, position = position_dodge(width = 0.5)) +
  coord_flip() +
  geom_hline(yintercept = 1, colour="grey", linetype = "longdash", size = 2) + 
  labs(x = "Ecosystem", y = "Number of Species/Number of Cameras/Day") + theme_classic()+ theme(legend.position = "none")

ggplot(summary_data, aes(ecosystem, capture_rate, color = factor(ecosystem))) +
  geom_point(size =2) +
  geom_errorbar(aes(ymin = capture_rate - var_caprate, ymax = capture_rate + var_caprate), size= 0.5, width=0.1, position = position_dodge(width = 0.5)) +
  coord_flip() +
  geom_hline(yintercept = 1, colour="grey", linetype = "longdash", size = 2) + 
  labs(x = "Ecosystem", y = "Number of Animals/Number of Cameras/Day") + theme_classic()+ theme(legend.position = "none")#variance is too small to see


ggplot(summary_data, aes(n_cams, capture_rate)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 2, color = "black", size = 2, se = FALSE)+ theme_classic()+ xlab("Number of Cameras")+ ylab ("Capture Rate")

ggplot(summary_data, aes(n_cams, richness_rate)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 2, color = "black", size = 2, se = FALSE)+ theme_classic()+ xlab("Number of Cameras")+ ylab ("Richness Rate")


#distribution of n_cam rates
ggplot(data_e1, aes(ecosystem, yi)) +
  geom_boxplot() +
  labs(x = "Ecoystem", y = "Number of Animals/Number of Cameras/Day") + 
  stat_summary(fun.y=mean, geom="point", shape=18, size = 5, color="black", fill="red") + theme_classic()

ggplot(data_e2, aes(ecosystem, yi)) +
  geom_boxplot() +
  labs(x = "Ecosystem", y = "Number of Species/Number of Cameras/Day") + 
  stat_summary(fun.y=mean, geom="point", shape=18, size = 5, color="black", fill="red") + theme_classic()


#cams as sampling effort
ggplot(data_e3, aes(days, yi)) +
  geom_point(aes(color = ecosystem)) + 
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 2, color = "black", se = FALSE) +
  labs(x = "Number of Days", y = "Number of Animals/Number of Cameras")+ theme_classic()+ labs(colour = "Ecosystem")

ggplot(data_e4, aes(days, yi, weight = vi)) +
  geom_point(aes(color = ecosystem)) + 
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 2, color = "black", size = 2, se = FALSE) +
  labs(x = "Number of Days", y = "Number of Species/Number of Cameras")+ theme_classic()+ labs(colour = "Ecosystem")


```


### Interpretations 
1. Net effect of increasing n_cam is positive.    
2. Ecosystem matters - all significantly positive except coniferous.
3. Increasing n_cams returns the highest rate of captures and of diversity in animals detected in grasslands.   
4. Increasing the number of days does not increase the capacity of the number of cams to detect more animals diversity.
5. Increasing the number of cameras only significanly increase capture rate in grassland. 

